Spatial distribution of environmental DNA in a nearshore marine habitat

### 2017-01-13 Manuscript/v1.1_PeerJ/response/response_to_reviews.tex

L90. Correct ``Actiniaria'' and ``Bryozoans''.
\begin{response}
  Thanks for catching these errors. We made these changes.\\
\end{response}

L91. Delete ``(Decapoda)''. These habitats hold many other Crustacea other than Decapoda as well. Actually, I would write ``barnacles (Sessilia) and other crustaceans'' in line 89.
\begin{response}
  Excellent suggestion; we made this change.\\
\end{response}

L93: ``which provide''
\begin{response}
  We made this change.\\
\end{response}

L94: The microscopic plankton is not composed just by ``diatoms and larvae''. Many important holoplanktonic organisms (e.g. copepods) could be mentioned here.
\begin{response}
  We did not mean to imply that the microscopic plankton is composed solely of diatoms and larve, we made a  statement and qualified it with two examples to give meaning and context: `a diverse assemblage of microscopic plankton including diatoms and larvae'.
	We mention diatoms because they are a widely-known, numerically abundant, and ecologically important group that falls outside the scope of this study on metazoans.
	We mention larvae because it subtly hints at our major finding that our samples captured many organisms best known for their sessile, benthic stages (e.g. barnacles).
	We added copepods to this list, though we recognize that it is necessarily incomplete.\\
\end{response}

@@

This is well designed study and we would like to congratulate the authors to this data set and its excellent analysis. However, we feel that the authors have made an important omission in the discussion.\\

Given that the main focus of this paper is to explore the spatial distribution of community eDNA we feel that the authors should consider the results of our study on lake fish communities (Hänfling et al. 2016). In our paper we have used extensive spatial sampling of three transects across our main study lake with a total of 66 individual samples and to our knowledge there is no other data set on macrobial eDNA metabarcoding with a comparable extent of explicit sampling. Our paper confirms that the spatial distribution of individual species reflects their abundance but also shows that the pattern of distribution is not random and ecologically meaningful. Although there might be important differences between freshwater and marine habitats we feel that the results from our study are extremely relevant here and should not be omitted especially when claiming novelty of discussion point 2.\\

Hänfling B, Lawson Handley L, Read DS, Hahn C, Li J, Nichols P, Blackman RC, Oliver A, Winfield IJ (2016) Environmental DNA metabarcoding of lake fish communities reflects long-term data from established survey methods. Molecular Ecology 25, 3101-3119.
\begin{response}
	% http://onlinelibrary.wiley.com/doi/10.1111/mec.13660/full
  Thank you for your comment and for drawing our attention to your paper.
	As you note therein, the rate of publication in this field can make it difficult to keep up.
	% Our omission was not intended as a slight, merely an oversight. % given the rapid pace of publication in this field.
	We commend your extensive environmental sampling; however, upon reading your paper, we have reservations about your interpretations of the results given the molecular protocols.
  \\

	We note that you conducted replicate PCRs of each environmental sample, and pooled them prior to sequencing.
	While this practice is not uncommon in the field, it precludes the ability to test for any effects of PCR on the communities.
	We believe this is particularly problematic in the case of your study because, as far as we understand your methods, you used indexed (`tagged') primers in a single PCR reaction.
	We have shown previously that this can bias datasets in such a way that completely obscures any underlying pattern (see: http://dx.doi.org/10.1371/journal.pone.0148698).
	As we report in that study, indexed primers can bias relative read abundance by as much as 70\%.
	That is, in one environmental sample, a species may comprise 70\% of the reads in PCR replicates amplified with one indexed primer set, while being completely absent from replicates amplified by another indexed primer set.
	This would very likely result in the detection of a significant difference by various statistical tests, such as the Chi-squared test you conducted to compare lake basins.
	% If we have misunderstood your methodology, or if you have results of replication tests on the indexed primers in community samples, we would be happy to consider further the results of your study.
  \\

### 2017-01-13 Manuscript/v1.1_PeerJ/carkeek_grid.tex
add Jameal's edits, render track changes version

Puget Sound is a deep, narrow fjord in Washington, USA, where a narrow band of shallow bottom hugs the shoreline and abruptly gives way to a central depth of up to 300 meters. This form allows the juxtaposition of communities associated with distinctly different habitats: shallow, intertidal benthos, and euphotic pelagic \citep{Burns1985}. At the upper reaches of the intertidal, the shoreline substrate varies from soft, fine sediment to cobble and boulder rubble. Soft intertidal sediments are inhabited by burrowing bivalves (Bivalvia), segmented worms (Annelida), and acorn worms (Enteropneusta), and in some lower intertidal and high subtidal ranges by eelgrass (\textit{Zostera marina}) \citep{Kozloff1973, Dethier2010} . Eelgrass meadows harbor epifaunal and infaunal biota, and attract transient species which use the meadows for shelter and to feed on resident organisms. Hard intertidal surfaces support a well-documented biota including barnacles (Sessilia) and other crustaceans, mussels (Bivalvia:Mytilidae), anemones (Actiniaria), sea stars (Asteroidea), urchins (Echinoidea), Bryozoans (Ectoprocta), crustaceans (Decapoda), and a variety of algae \citep{Dethier2010}. Hard bottoms of the lower intertidal and high subtidal are home to macroalgae such as Laminariales and Desmarestiales which provide habitat for a distinct community of fish and invertebrates. The upper pelagic is home to a diverse assemblage of microscopic plankton including diatoms and larvae \citep{Strickland1983}, as well as transitory fish and marine mammals.

>>>>

Puget Sound is a deep, narrow fjord in Washington, USA, where a narrow band of shallow bottom hugs the shoreline and abruptly gives way to a central depth of up to 300 meters. This form allows the juxtaposition of communities associated with distinctly different habitats: shallow, intertidal benthos, and euphotic pelagic \citep{Burns1985}. At the upper reaches of the intertidal, the shoreline substrate varies from soft, fine sediment to cobble and boulder rubble. Soft intertidal sediments are inhabited by burrowing bivalves (Bivalvia), segmented worms (Annelida), and acorn worms (Enteropneusta), and in some lower intertidal and high subtidal ranges by eelgrass (\textit{Zostera marina}) \citep{Kozloff1973, Dethier2010} . Eelgrass meadows harbor epifaunal and infaunal biota, and attract transient species which use the meadows for shelter and to feed on resident organisms. Hard intertidal surfaces support a well-documented biota including barnacles (Sessilia) and other crustaceans, mussels (Bivalvia:Mytilidae), anemones (Actiniaria), sea stars (Asteroidea), urchins (Echinoidea), Bryozoans (Ectoprocta), crustaceans (Decapoda), and a variety of algae \citep{Dethier2010}. Hard bottoms of the lower intertidal and high subtidal are home to macroalgae such as Laminariales and Desmarestiales which provide habitat for a distinct community of fish and invertebrates. The upper pelagic is home to a diverse assemblage of microscopic plankton including diatoms, copepods, and larvae \citep{Strickland1983}, as well as transitory fish and marine mammals.
